---
title: "Solution Assignment Week 3"
author: "GSS Data Analysis Course"
output: html_document
---

```{r setup, include=TRUE, echo = TRUE, warning=FALSE, message = FALSE}
# Setup
library(tidyverse)
library(readxl)
library(writexl)
```

## Part 1 - R Practice:
1. **Import the population data.**
```{r}
# Importing population data
population <- read_xlsx("input/population.xlsx") 

# Display first five rows
head(population,5)
```

2. **Filter the dataset:**
   - Include only rows where the region is one of the following: Western, Greater Accra, Volta, Savannah, Northern, and Ashanti.
```{r}
# Subsetting the data to include the specified regions
pop_six_reg <- population %>% 
  filter(region %in% c("Western","Greater Accra", "Volta", "Savannah", "Northern", "Ashanti"))

# Checking the unique values in the region columns after subsetting
unique(pop_six_reg$region)
```
Alternatively, you can define the vector outside the piped operation. 
```{r}
# Creating a vector of the six regions
six_reg <- c("Western","Greater Accra", 
                       "Volta", "Savannah", 
                       "Northern", "Ashanti")

# Subset data to include only the six regions
pop_six_reg_alt <- population %>% 
  filter(region %in% six_reg)

# Display the unique regions
unique(pop_six_reg_alt$region)
```
Note that the %in% is used instead of == in filtering when you want to extract a subset of data based on whether certain elements are present in a given vector.

3. **Select columns:**
   - Select the columns region, a11d, and p02 from the dataset.
```{r}
# Subset data to include only region, a11d, and p02
selected_data <- population %>% 
  select(region, a11d, p02)

# Checking the columns within the selected data
colnames(selected_data)
```

4. **Rename columns:**
   - Rename the column urbrur to urban_rural, a11d to sex, and p02 to age.
```{r}
# Rename the column names
renamed_data <- population %>% 
  rename(urban_rural = urbrur,
         sex = a11d,
         age = p02)

# Display the column names (you can either use names() or colnames())
names(renamed_data)
```

5. **Group data and calculate mean age:**
   - Group the data by region and calculate the mean age for each region using the `group_by` and `summarise` functions.
```{r}
# Finding the mean age by region
mean_age <- population %>% 
  mutate(age = as.numeric(p02)) %>% 
  group_by(region) %>% 
  summarise(mean_age = mean(age, na.rm = TRUE))

# Print the mean age dataframe
mean_age
```
Remember, we had to convert the age column to numeric because it was loaded as a character. You can use the function typeof() or class() to confirm before converting.

6. **Pivot data:**
   - Pivot the data to make a wider format with region as columns and mean age as values.
```{r}
# Pivot the data into a wide format
wider_format <- mean_age %>% 
  pivot_wider(names_from = region, 
              values_from = mean_age)

# Print the wide format
wider_format
```

### Feel free to reach out to the facilitators with questions!
