---
title: "Solution Assignment Week 4"
author: "GSS Data Analysis Course"
output: html_document
---

```{r setup, include=TRUE, echo = TRUE, warning=FALSE, message = FALSE}
# Setup
library(tidyverse)
library(lubridate)
library(tseries)
library(dplyr)
library(scales)
```

## Part 1 - R Practice:
1. **Make a bar graph for average miles per gallon (mpg) by number of cylinders (cyl):**
   - Using the base R `mtcars` dataset, create a bar graph showing the average miles per gallon for each unique number of cylinders (cyl).
```{r}
#Loading data
data(mtcars)

# Inspect data
str(mtcars)

# Generate data for the graph: Calculating the avg_mpg per unique number of cylinders
av_mpg <- mtcars %>% 
  group_by(cyl) %>% 
  summarise(avg_mpg = mean(mpg))

# Print the data we will use for the graph
av_mpg

# Creating a bar graph  using the base package in R
barplot(av_mpg$avg_mpg, names.arg = av_mpg$cyl, 
        main = "Average MPG by number of cylinders",
        xlab = "Number of cylinders", ylab = "Average MPG",
        col = "salmon", border = "black")

# Now create the same using ggplot
# ggplot is much easier to use for nice visualizations!
av_mpg %>% 
  ggplot(aes(x = as.factor(cyl), y = avg_mpg)) +
  geom_bar(stat = "identity", fill = "salmon", color = "black") +
  labs(title = "Average MPG by number of cylinders",
       x = "Number of cylinders", y = "Average MPG") +
  theme_minimal()
```

2. **Make a scatter plot for petal length vs petal width with species coloring:**
   - Use the `iris` dataset in base R to create a scatter plot with petal length on the x-axis and petal width on the y-axis. Color the points based on the species.
```{r}
# Load the data
data(iris)

# Create plot
iris %>% 
  ggplot(aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  geom_point() +
  labs(title = "Scatter Plot of Petal Length vs. Petal Width",
       x = "Petal Length", y = "Petal Width") +
  theme_classic()

```

3. **Make a line graph for unemployment and population trends over time:**
   - Using the `economics` dataset, create a line graph with time on the x-axis and two lines representing the trends of unemployment and population over time. Differentiate the lines by color.
```{r}
# Load data
data("economics")

# Inspect the data
str(economics)

# Create graph
economics %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = unemploy, color = "Unemployment"), size = 1) +
  geom_line(aes(y = pop, color = "Population"), size = 1) +
  labs(title = "Unemployment and Population Trends Over Time",
       x = "Year", y = "Number") +
  scale_color_manual(values = c("Unemployment" = "blue", "Population" = "green")) +
  scale_y_continuous(labels = comma, 
                     limits = c(0, max(economics$pop))) +  
  theme_minimal()
```

4. **Make a box plot for diamond prices distribution by cut, color, and clarity:**
   - Using the `diamonds` dataset, generate a box plot showing the distribution of diamond prices for each combination of cut and color. Use facet grid to separate the plots based on the diamond's clarity.
```{r}
# Load the data
data("diamonds")

# Make the graph
diamonds %>% 
  ggplot(aes(x = cut, y = price, fill = color)) +
  geom_boxplot() +
  theme_light()

# Now add the facet grid
diamonds %>% 
  ggplot(aes(x = cut, y = price, fill = color)) +
  geom_boxplot() +
  facet_grid(clarity ~ ., scales = "free_y", space = "free_y", switch = "y") +
  labs(title = "Distribution of Diamond Prices by Cut, Color, and Clarity", x = "Cut", y = "Price") +
  scale_fill_brewer(palette = "Set3") +
  theme_minimal()
```

### Feel free to reach out to the facilitators with questions!
